### 2016-03-29 chapters/concept.tex
conclusion added, references included
Added reviews part 1/2

This chapter describes the general workflow we aim to archive within this work. This workflow can be seen in Figure~\ref{img-workflow}. To start off, some kind of software product line is required. It's source code has to be structured in a way that allows for an algorithm to recognize the individual features as well as their interrelations, especially the hierarchical dependencies.

@@

The two major steps of extraction of the feature model and the creation and usage of the questionnaire are described in detail in the following sections. As a final result of these steps stands a product targeting the users needs. Supporting the process towards that result is what we aimed at with this work.

\subsection{Extraction of a Feature Model}
Out of the given structure of the source code base we extract the hierarchies of the features as well as additional dependencies. These are the input for the automated generation of a feature model. This allows for overview over the product line and it's variability. During our efforts to implement an automated generation of a feature model we encounter some problems for which we have to find solutions/workarounds.

The first problem consists of finding the structures of the given source code and processing them algorithmically. Most software complies to conventions regarding the naming and structuring of directories. We try to make use of these conventions. But as they are just conventions and not rules, in most source code one will find violations of the conventions. Instead of interrupting the whole process, we implement error handling. The result is a notification for the user to alert the developers about the error. Also, the remaining information, which aren't immediately affected by the error, are still extracted and used for the generation of the feature model, if possible.

@@

Another error we found is one ore more features having a parent feature which doesn't exist as a code artifact itself. Thus, the parent feature doesn't exists in the feature model and it's child features can't be placed in the hierarchy in the correct place. To still include these features in the feature model we create an abstract feature for the parent feature. The abstract feature is a direct child of the root feature. But as it doesn't have any describing code, it only has a name and no additional descriptive details.

In each step of the creation of the feature model, e.g. after adding a feature or constraint to the feature model, it is checked by a SAT-solver to avoid creating an invalid feature model which has no valid configurations (see section~\ref{ch:sat}). The extracted feature model is also used in the following step of configuration. There, it visualizes the interrelations of the features and supports the understanding of the configuration steps.

@@

This work therefore introduces a method to allow experts to apply their knowledge and understanding to a whole product line during the development. Thus, users are enabled to draw on this knowledge whenever a configuration is taking place.

In this work, we made the decision to use a questionnaire-based approach. In the progress of implementation, experts also design a questionnaire. They do so in such a way, that a user has to answer a given amount of questions to perform the configuration of a product. Depending on the implementation of the questionnaire, a partial or even complete configuration can be archived by a user through just answering the questions of the questionnaire.

The selection of features gets lifted to a higher level of abstraction. The user only has to decide between the possible answers presented to him in the questionnaire to best fit to his use-case. Internally the selected answers are mapped to a specified (un-)selection of features so the user avoids the hassle of considering implementation-details of each feature. This effectively redesigns the process of configuration in such a way, that a user is independent of the knowledge about the details of the implementation and can focus on tailoring the product line to his specific use-case.

This highly depends on the implementation of the questionnaire during development. Our work therefore introduces a set of tools to easily integrate such a questionnaire. The following paragraphs will give an overview over the possible definitions of a questionnaire.

Each page of the questionnaire is defined independently. It always contains at least a Question and more then one possible answer. The answers can be grouped analogous to the grouping of features in a feature model: \textit{or} (at least one answer has to be selected), \textit{alternative} (exactly one answer has to be selected) and \textit{and} (any number of answers can be selected).

Each answer internally has a mapping to a set of features. This set of features defines which features are selected or specifically unselected in the case of that answer being chosen by the user.

Each answer can also have an indicator to define which page of the questionnaire is to be displayed next. An answer can also indicate the end of the questionnaire. If no next page is defined the questionnaire will continue with the next page within it's definition. This allows a dynamic conditional design of the questionnaire so the user is only confronted with the exact set of questions needed to configure the variant for his specific use-case. This also allows the user to skip questions or cancel the configuration before finishing it and thus creating a partial configuration.

We also introduce a data structure to hold the definition of a questionnaire. To archive easy integration we decided for a definition in XML. We defined the necessary tags to create a questionnaire which are displayed in figure~\ref{fig:xmlFigure}.

>>>>

This chapter describes the general workflow we propose within this work. This workflow can be seen in Figure~\ref{img-workflow}. To start off, some kind of software product line is required. It's source code has to be structured in a way that allows for an algorithm to recognize the individual features as well as their interrelations, as these can be very costly to implement by hand.

{\color{red} Pfeil von Fragebogen zu Generator sollte zu der Selektion zeigen.}

@@

The two major steps of extraction of the feature model and the creation and usage of the questionnaire are described in detail in the following sections. As a final result of these steps stands a product targeting the users needs.

\subsection{Extraction of a Feature Model}
Out of the given structure of the source code base we extract the hierarchies of the features as well as additional dependencies. These are the input for the automated generation of a feature model. This allows for overview over the product line and it's variability. During our efforts to implement an automated generation of a feature model we encounter some problems for which we have to find solutions or workarounds.

The first challenge consists of finding the structures of the given source code and processing them programmatically. Most software complies to conventions regarding the naming and structuring of directories. We try to make use of these conventions. However, in most source code there are violations within their own conventions. Instead of interrupting the whole process, we implement error handling. The result is a notification for the user to alert the developers about the error. Also, the remaining information, which aren't immediately affected by the error, are still extracted and used for the generation of the feature model, if possible.

@@

Another error we found is one or more features having a parent feature which doesn't exist as a code artifact itself. Thus, the parent feature doesn't exists in the feature model and it's child features can't be placed in the hierarchy in the correct place. To still include these features in the feature model we create an {\color{red}abstract feature (Zitat)} for the parent feature. The abstract feature is a direct child of the root feature. As it doesn't have any describing code, it only has a name and no additional descriptive details.

In each step of the creation of the feature model, e.g. after adding a feature or constraint to the feature model, it is checked by a SAT-solver to avoid creating an invalid feature model (represented by a void object) which has no valid configurations (see section~\ref{ch:sat}). The extracted feature model is also used in the following step of configuration. There, it visualizes the interrelations of the features and supports the understanding of the configuration steps.

@@

This work introduces a method allowing experts to apply their knowledge and understanding to a whole product line during the development. Whenever a variation is configured using the questionnaire, the user benefits from the invested domain knowledge.

In this work, we made the decision to use a questionnaire-based approach \cite{qbvm}. In the progress of implementation, experts also design a questionnaire. They do so in such a way, that a user has to answer a given amount of questions to perform the configuration of a product. Depending on the implementation of the questionnaire, a partial or even complete configuration can be archived by a user through just answering the questions of the questionnaire.

The selection of features gets lifted to a higher level of abstraction. The user only has to decide between the possible answers presented to him in the questionnaire to best fit to his use-case. Internally the selected answers are mapped to a specified (de-)selection of features. This way the user avoids the hassle of considering implementation-details of each feature. This effectively redesigns the process of configuration in such a way, that a user is independent of the knowledge about the details of the implementation and can focus on tailoring the product line to his specific use-case.

\subsection{Questionnaire Implementation}

The quality of the resulting variation highly depends on the implementation of the questionnaire during development. Our work therefore introduces a set of tools to easily integrate such a questionnaire. The following paragraphs will give an overview over the possible definitions of a questionnaire.

{\color{red} TODO: add screenshots}

Each page of the questionnaire is defined independently. It always contains at least a question and more than one possible answer. The answers can be grouped analogous to the grouping of features in a feature model: \textit{or} (at least one answer has to be selected), \textit{alternative} (exactly one answer has to be selected) and \textit{and} (any number of answers can be selected).

Each answer internally has a mapping to a set of features. In the case a specific answer is chosen, the corresponding features are selected or deselected as specified.

Each answer can also have an indicator to define which page of the questionnaire should be displayed next. An answer can also indicate the end of the questionnaire. If no next page is defined the questionnaire will continue with the next page within it's definition. This allows a dynamic conditional design of the questionnaire so the user is only confronted with the exact set of questions needed to configure the variant for his specific use-case. This also allows the user to skip questions or cancel the configuration before finishing it and thus creating a partial configuration.

We also introduce a data structure to hold the definition of a questionnaire. To archive easy integration we decided for a definition in XML. We defined the necessary tags to create a questionnaire which are displayed in Figure~\ref{fig:xmlFigure}.

### 2016-03-29 chapters/conclusion.tex
Added reviews part 1/2

In this work we aim at finding a way to support the process of configuring an existing software product line to generate an actual product.

To do so, we firstly use feature models as formalism to structure the existing code artifacts. The feature model gets filled with the information we get from analyzing the structure and conventions of the existing source code. The resulting feature model contains every code artifact as a feature and also includes the constraints and restrictions.

We also introduce a questionnaire to support the configuration of a product from the given product line. A questionnaire needs to be designed carefully by the developers or experts on the domain of the product line. It allows a mapping between answers to simple questions and actual features of the product line. With this method, a user doesn't have to undergo the process of configuration in the sense of (un-)selecting a specific set of features. Instead, he only has to answer questions according to his use case to do the configuration of his specific product. As the tool uses the underlying FeatureIDE, it also makes use of its functions, e.g. validity checks. This additionally supports the usability and correctness of the whole process.

@@

To make meaningful use of the tools introduced in this work, one would at least need a software which is structured in a ``feature-like'' manner. This structure has to be in a way that allows a feature model to be extracted from it. Alternatively, this first step can be skipped, if a feature model already exists for the given software product line.

Also, the features have to have some kind of interrelations that allows abstract questions to map to them. If that is not the case, the questionnaire degenerates to a simple (un-) selection of each feature.

>>>>

{\color{red}Ihr geht hier in die falsche Richtung. Schreibt einfach was jetzt geht, ihr bracht hier nicht zu diskutieren was notwendig ist oder so (wie z.b. im letzten absatz)}

In this work we propose a questionnaire-based approach to support the process of configuring an existing software product line to generate an actual product. To do so, we use feature models as formalism to structure the existing code artifacts. The feature model gets filled with the information we get from analyzing the structure and conventions of the existing source code. The resulting feature model contains every code artifact as a feature and also includes the constraints and restrictions.

We also introduce a questionnaire to support the configuration of a product from the given product line. A questionnaire needs to be designed carefully by the developers or experts on the domain of the product line. It allows a mapping between answers to simple questions and actual features of the product line. With this method, a user doesn't have to undergo the process of configuration in the sense of \mbox{(de-)selecting} a specific set of features. Instead, the user only has to answer questions according to its use case to do the configuration of his specific product. As the tool uses the underlying FeatureIDE, it also makes use of its functions, such as validity checks. This additionally supports the usability and correctness of the whole process.

@@

To make meaningful use of the tools introduced in this work, a developer would at least need a software which is structured in a ``feature-like'' manner. This structure has to be in a way that allows a feature model to be extracted from it. Alternatively, this first step can be skipped, if a feature model already exists for the given software product line.

Also, the features need to have some kind of interrelations that allows abstract questions to map to them. If that is not the case, the questionnaire degenerates to a simple (un-) selection of each feature.

### 2016-03-29 chapters/scenarios.tex
todos are now in the pdf
Added reviews part 1/2

In the fallowing chapter the workflow described in Chapter~\ref{ch:workflow} is carried out based on the current project status of \textit{Odoo}. At first a feature model is generated based on the naming conventions of the project folders and the configuration files contained therein. Afterwards the feature model is corrected from any errors that occurred during the generation process. The resulting model combined with domain knowledge allows for the creation of a questionnaire that takes a minimum of choices to configure a custom tailored application.

All the Odoo features lie inside of a folder named "addons". The code of each feature is contained in its own folder which is named like the parent feature with a suffix of its own name. According to this naming convention the feature "website\textunderscore sale" is a child of the feature "website", and parent to "website\textunderscore sale\textunderscore stock". In addition to the naming convention, there is a configuration file, written in python in each of the feature folders. These files contain a description of the feature, a name, that doesn't always match with the folder name and the dependencies that are required for this feature to work.
By first creating a list of all features based on the folder names, one can match them to the names inside of the configuration files. At this point we realize, that there are some conflicts in the naming of the 261 features. For example the feature "sale" is a standalone feature and has nothing to do with the sale from the feature "event\textunderscore sale". As the default FeatureIDE doesn't support multiple features with the same name, we decided to leave the full names with the underscores. In addition, there are some folder names where the underscore doesn't separate features, like "point\textunderscore of\textunderscore sale". For those we implemented naming exceptions where the user can add strings that represent a single feature.

@@

%TODO: Auswertung
TODO:

Also state how some of the questions and their mapping in the questionnaire were developed.

Finally, show the process of configuration through running a fictinal user over the questionnaire and document his decisions as well as the resulting configurations.

>>>>

{\color{red} erstmal beschreiben was odoo ist und warum ihr euren ansatz darauf anwenden wollt und könnt dann bitte auch noch erklären wie das projekt strukturiert is, wie groß es ist, wie viele module es gibt etc.referenz zur odoo webseite oder repository}

In the fallowing chapter the workflow described in chapter~\ref{ch:workflow} is carried out based on the current project status of \textit{Odoo}. At first, a feature model is generated based on the naming conventions of the project folders and the configuration files contained therein. Afterwards the feature model is corrected from any errors that occurred during the generation process. The resulting model combined with domain knowledge allows for the creation of a questionnaire that takes a minimum of choices to configure a custom tailored application.

{\color{red} bitte die komplette struktur vorab klären}

All the Odoo features lie inside of a folder named "addons". The code of each feature is contained in its own folder which is named like the parent feature with a suffix of its own name. According to this naming convention the feature "website\textunderscore sale" is a child of the feature "website", and parent to "website\textunderscore sale\textunderscore stock". In addition to the naming convention, there is a configuration file, written in python in each of the feature folders. These files contain a description of the feature, a name, that doesn't always match with the folder name and the dependencies that are required for this feature to work.
By first creating a list of all features based on the folder names, we can match them to the names inside of the configuration files. At this point we realize, that there are some conflicts in the naming of the 261 features. For example the feature "sale" is a standalone feature and has nothing to do with the sale from the feature "event\textunderscore sale". As the default FeatureIDE doesn't support multiple features with the same name, we decided to leave the full names with the underscores. In addition, there are some folder names where the underscore doesn't separate features, like "point\textunderscore of\textunderscore sale". For those we implemented naming exceptions where the user can add strings that represent a single feature.

@@

{\color{red}TODO: Fragebogen}

{\color{red}TODO: Auswertung}

%TODO: Auswertung

%Also state how some of the questions and their mapping in the questionnaire were developed.

%Finally, show the process of configuration through running a fictinal user over the questionnaire and document his decisions as well as the resulting configurations.
